Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)

### 2021-06-21 doc/manuscript/ESSDmanuscript.Rmd
litter/ O.horizon/ organic.layer distinction
https://github.com/forc-db/ForC/issues/240

#19

Litter is defined as including " all  non-living  biomass  with  a  diameter  less  than  a  minimum  diameter  chosen  by  the  country  (for  example 10 cm), lying dead, in various states of decomposition above the mineral or organic soil. This includes litter,  fumic,  and  humic  layers.  Live  fine  roots  (of  less  than  the  suggested  diameter  limit for  belowground  biomass) are included in litter where they cannot be distinguished from it empirically." (2003 IPCc GPG for LULUCF (https://www.ipcc-nggip.iges.or.jp/public/gpglulucf/gpglulucf_files/Glossary_Acronyms_BasicInfo/Glossary.pdf)

@@

We clarified the definition of *organic.layer* (O-horizon, including *litter*) and added the variable *litter* (relatively undecomposed plant material/ OL horizon).

>>>>

Litter is defined as including " all  non-living  biomass  with  a  diameter  less  than  a  minimum  diameter  chosen  by  the  country  (for  example 10 cm), lying dead, in various states of decomposition above the mineral or organic soil. This includes litter (OL),  fumic (OF),  and  humic (OH) layers.  Live  fine  roots  (of  less  than  the  suggested  diameter  limit for  belowground  biomass) are included in litter where they cannot be distinguished from it empirically." (2003 IPCc GPG for LULUCF (https://www.ipcc-nggip.iges.or.jp/public/gpglulucf/gpglulucf_files/Glossary_Acronyms_BasicInfo/Glossary.pdf)

@@

To provide better definition of the previously existing variable *organic.layer*, which has a nebulous definition that reflects the varied definitions adopted by original studies, we added two clearly defined variables: *litter* (relatively undecomposed plant material/ OL horizon), and *O.horizon* (entire O-horizon, including *litter* (OL)).

### 2021-06-21 doc/manuscript/ESSDmanuscript.tex
add a bit of text
litter/ O.horizon/ organic.layer distinction
https://github.com/forc-db/ForC/issues/240

#19

Litter is defined as including " all non-living biomass with a diameter
less than a minimum diameter chosen by the country (for example 10 cm),
lying dead, in various states of decomposition above the mineral or
organic soil. This includes litter, fumic, and humic layers. Live fine
roots (of less than the suggested diameter limit for belowground
biomass) are included in litter where they cannot be distinguished from
it empirically." (2003 IPCc GPG for LULUCF
(https://www.ipcc-nggip.iges.or.jp/public/gpglulucf/gpglulucf\_files/Glossary\_Acronyms\_BasicInfo/Glossary.pdf)

@@

@@

To support export of data to EFDB, and to improve the overall quality of
the ForC database, we added ten increment variables, implemented some
modest restructuring, resolved duplicate records, and conducted quality
control. This section describes changes relative to ForC v2.0
\citep{anderson-teixeira_forc_2018}.

\subsection{Increment variables}

We added ten increment variables to the set of named and defined
variables (or 20, counting \_OM and \_C versions), which previously
included only one (aboveground biomass increment, \emph{delta.agb}).
\emph{(https://github.com/forc-db/IPCC-EFDB-integration/issues/6)} These
are directly related to C stocks as previously defined in ForC, with
``\emph{delta.}'' added in front of the variable name.

@@

>>>>

Dead organic matter includes all non-living biomass that is not within
the mineral soil layer and smaller than the litter size threshold.

Dead wood is defined as\ldots{}

Litter is defined as including " all non-living biomass with a diameter
less than a minimum diameter chosen by the country (for example 10 cm),
lying dead, in various states of decomposition above the mineral or
organic soil. This includes litter (OL), fumic (OF), and humic (OH)
layers. Live fine roots (of less than the suggested diameter limit for
belowground biomass) are included in litter where they cannot be
distinguished from it empirically." (2003 IPCc GPG for LULUCF
(https://www.ipcc-nggip.iges.or.jp/public/gpglulucf/gpglulucf\_files/Glossary\_Acronyms\_BasicInfo/Glossary.pdf)

@@

Soil organic matter is defined as\ldots{}

@@

To support export of data to EFDB, and to improve the overall quality of
the ForC database, we defined \#\# new variables, implemented some
modest restructuring, resolved duplicate records, and conducted quality
control. This section describes changes relative to ForC v2.0
\citep{anderson-teixeira_forc_2018}.

\subsection{Defining new variables}

We added eleven increment variables to the set of named and defined
variables (or 22, counting \_OM and \_C versions), which previously
included only one (aboveground biomass increment, \emph{delta.agb}).
\emph{(https://github.com/forc-db/IPCC-EFDB-integration/issues/6)} These
are directly related to C stocks as previously defined in ForC, with
``\emph{delta.}'' added in front of the variable name.

@@

To provide better definition of the previously existing variable
\emph{organic.layer}, which has a nebulous definition that reflects the
varied definitions adopted by original studies, we added two clearly
defined variables: \emph{litter} (relatively undecomposed plant
material/ OL horizon), and \emph{O.horizon} (entire O-horizon, including
\emph{litter} (OL)).
